
# Regression analysis

This week I learned to do regression diagnostics with R and learned to use R-markdown a little more. 

Data comes from international survey of Approaches to Learning.


## Exercise 1: structure of data

```{r}
student_2014<-read.table("http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/learning2014.txt", header=TRUE, sep=",")
head(student_2014)
dim(student_2014)
str(student_2014)
```

The data includes 7 variables and 166 students. Variables include for example gender, age, attitude towards statistics and exam points.


## Exercise 2: more about structure of data

```{r}
summary(student_2014)
hist(student_2014$age)
hist(student_2014$attitude)
hist(student_2014$deep)
hist(student_2014$stra)
hist(student_2014$surf)
hist(student_2014$points)
pairs(student_2014)
```

Most students seem to be female and quite young. Points and attitudes seem to be quite normally distributed as the variables that measure deep-/strategic- and surface approach.


## Exercise 3: regression analysis

```{r}
fit<-lm(points~attitude+age+deep,data=student_2014)
fit
summary(fit)
```

The only variable that has a statistically significant relation with exam points is the attitude-variable. Age and deep approach do not have a statistically significant relationship with exam points.

```{r}
fit2<-lm(points~attitude, data=student_2014)
fit2
summary(fit2)
```


## Exercise 4: interpreting model parameters

The only explanatory varialbe, attitude, is a significant predictor of exam points (p -value close to 0.) When attitude increases one point, exam points increases on average 3.5255 points. Multiple R-square is 0.1906, meaning that the attitude variable can explain about one fifth of the variance in exam points, which is fairly good.


## Exercise 5: assumptions of regression model

```{r}
par(mfrow=c(2,2))
plot(fit2)
```

When looking at fitted values vs. residuals one can see whether there are non-linear patterns. In this case there are not. When looking at normal Q-Q -plot we can see that whether residuals are normally distributed. In this case they are. Scale location -plot shows that residuals are also spread equally along the ranges of predictors. Residuals vs. leverage -plot shows whether there are influential cases against the regression line. In this case there are not. 

So, we can say that regression model is a suitable method with this data.
